---
title: "Take-home Exercise 2 – Clustering Module Prototype"
author: "Chun-Han"
date: "February 28, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Overview

This take-home exercise develops a prototype of the Customer Segmentation Module for the final Shiny application.

The dataset is already structured at the customer level, meaning each row represents a unique customer with demographic, behavioural, engagement, and satisfaction attributes. Therefore, no transaction-level aggregation is required.

The purpose of this prototype is to:

-   Feature engineering from transaction-level data

-   Feasibility of k-means clustering

-   Interactive parameters (k and variable selection)

-   Visual outputs for cluster exploration

## 1 Getting Started

### 1.1 Installing and loading the packages

Before performing analysis, the required packages are loaded. These packages support data manipulation, clustering evaluation, and visualisation.

| Library | Description |
|------------------------------------|------------------------------------|
| **pacman** | Install and load multiple packages in one command. |
| **tidyverse** | Data wrangling + ggplot2 visualization workflow . |
| **lubridate** | Date parsing and date arithmetic. |
| **cluster** | Clustering evaluation utilities (e.g., silhouette). |
| **factoextra** | Helper functions for visualising clusters and PCA results . |
| **DT** | Interactive tables for cluster profile summaries in Shiny. |
| **shiny** | UI and server framework for integrating this module into the final Shiny app. |
| **plotly** | Interactive plotting for cluster maps and exploratory visuals. |

```{r}
pacman::p_load(
  pacman, tidyverse, lubridate, cluster, factoextra, DT, shiny, plotly
)
```

## 2 Data Preparation

### 2.1 Loading the Dataset

We first import the customer-level dataset and inspect its structure to confirm variable types.

```{r}
customers <- read_csv("data/customer_data.csv")
glimpse(customers)
```

### 2.2 Data Integrity Checks

#### 2.2.1 Missing Values

Missing values may distort clustering results. Therefore, we inspect NA counts.

```{r}
customers %>%
  summarise(across(everything(), ~sum(is.na(.))))
```

The missing value inspection shows that all 54 variables contain zero missing entries.

This indicates that the dataset is fully observed at the customer level. 
No imputation or row removal procedures are required.

#### 2.2.2 Duplicate Records

Since clustering is performed at the customer level, duplicate rows would bias the segmentation.

```{r}
sum(duplicated(customers))
```

The result indicates that no duplicate rows exist, confirming dataset integrity.

## 3 Variable Selection for Clustering

Clustering should rely on behavioural and engagement signals, rather than identifiers or pre-existing labels.

The following variables are selected:

```{r}
cluster_vars <- c(
  "monthly_transaction_count",
  "average_transaction_value",
  "total_transaction_volume",
  "transaction_frequency",
  "avg_daily_transactions",
  "app_logins_frequency",
  "feature_usage_diversity",
  "active_products",
  "credit_utilization_ratio",
  "weekend_transaction_ratio",
  "customer_tenure"
)
```

These variables capture:

- Transaction intensity

- Spending behaviour

- App engagement

- Product breadth

- Credit behaviour

- Customer tenure

## 4 Preprocessing

Clustering algorithms such as K-means rely on Euclidean distance.
Therefore, proper preprocessing is essential to:

- Prevent scale dominance

- Reduce skewness effects

- Ensure fair distance computation

The preprocessing pipeline follows three steps:

- Preserve original data for interpretation

- Apply log transformation to skewed variables

- Standardise variables before clustering

### 4.1 Create Original Dataset 

Before transformation, we preserve the original numeric values. This dataset will later be used to compute cluster profile summaries.

```{r}
df_original <- customers %>%
  select(all_of(cluster_vars)) %>%
  mutate(across(everything(), as.numeric)) %>%
  drop_na()

dim(df_original)
```

### 4.2 Skewness Assessment

Before applying log transformation, we quantify skewness to determine whether transformation is necessary.

```{r}
library(e1071)

skew_table <- df_original %>%
  summarise(across(everything(), ~skewness(.))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "skewness") %>%
  arrange(desc(abs(skewness)))

skew_table
```

### 4.3 Interpretation

Variables with absolute `skewness > 1` are considered highly skewed and candidates for log transformation.

```{r}
skew_vars <- skew_table %>%
  filter(abs(skewness) > 1) %>%
  pull(variable)

skew_vars
```

### 4.4 Log Transformation

Since the skewness assessment identified highly skewed variables,
we apply log transformation only to those variables.

We use `log1p(x)` (log(1 + x)) to safely handle zero values.

```{r}
df_model <- df_original %>%
  mutate(across(all_of(skew_vars), ~log1p(.)))
```

### 4.5 Skewness Comparison

To validate the effectiveness of transformation, we recompute skewness after applying log transformation.

```{r}
skew_after <- df_model %>%
  summarise(across(everything(), ~skewness(.))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "skew_after")

skew_compare <- skew_table %>%
  left_join(skew_after, by = "variable")

skew_compare
```

This confirms that transformation improved distribution symmetry
without over-transforming already well-behaved variables.

### 4.6 Standardisation

K-means clustering relies on Euclidean distance, which is sensitive to variable scale.

Without standardisation:

- Variables with larger magnitude (e.g., transaction volume)
would dominate distance computation.

- Smaller-scale variables (e.g., ratios) would have negligible influence.

Therefore, we apply z-score scaling.

```{r}
df_scaled <- scale(df_model)
```

## 5 Clustering Implementation

With preprocessing completed, we proceed to implement clustering using K-means.

K-means is selected because:

- It is computationally efficient

- It is suitable for numeric, standardised variables

- It integrates well with interactive Shiny applications

- It allows user-controlled selection of `k`

### 5.1 Determining Optimal Number of Clusters

Before fixing k, we evaluate cluster compactness using the Elbow Method.

The Elbow Method examines the total within-cluster sum of squares (WSS) as k increases.

::: panel-tabset
### Plot

```{r}
#| echo: false

set.seed(2022)

wss <- map_dbl(2:15, function(k){
  kmeans(df_scaled, centers = k, nstart = 20)$tot.withinss
})

elbow_df <- tibble(
  k = 2:15,
  wss = wss
)

ggplot(elbow_df, aes(k, wss)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Elbow Method",
    x = "Number of Clusters (k)",
    y = "Total Within-Cluster Sum of Squares"
  )
```

### Code

```{r}
#| eval: false

set.seed(2022)

wss <- map_dbl(2:15, function(k){
  kmeans(df_scaled, centers = k, nstart = 20)$tot.withinss
})

elbow_df <- tibble(
  k = 2:15,
  wss = wss
)

ggplot(elbow_df, aes(k, wss)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Elbow Method",
    x = "Number of Clusters (k)",
    y = "Total Within-Cluster Sum of Squares"
  )
```
:::

